\clearpage
\section{Method}

\subsection{Participants}

The dataset~\parencite[]{uci94census} for this study originates from the 
1994 U.S. Census, which 
involved a diverse range of participants to provide a comprehensive snapshot 
of the U.S. population. This extensive coverage makes the findings relevant to 
the adult population of the United States during that period.

The data captures key demographic details critical for examining income 
disparities, including age, gender, race/ethnicity, education level, 
marital status, and employment status. 
Specifically, it categorizes gender into male or female; race/ethnicity 
includes White, Black, Asian, Hispanic, among others; and education 
levels vary from none to advanced degrees.

The census participants were not specifically recruited like in targeted 
research studies but were included through a national initiative aimed at 
documenting the demographic attributes of the U.S. population. Participation
was solicited from households nationwide, covering both urban and rural areas
to ensure a representative sample.

Although the census gathered data from millions, this study analyzes a subset
of 32,561 complete records, chosen for their comprehensive demographic details
needed to study income variations. No power analysis was conducted as the 
census was designed to encompass as broad a population segment as possible, not 
to meet specific sample size calculations for statistical power.

The data was gathered specifically to analyze the earned income of the 
population, and the collection was conducted according to the following 
criteria:
\begin{description}
\item[AAGE > 16] Keep only working-age adults;
\item[AGI > 100] Adjusted Gross Income;
\item[AFNLWGT > 1] Final Weight meaning the number of people that is believed
                   that this entry represents;
\item[HRSWK > 0] Participants with working hours reported.
\end{description}

The dataset comes in comma-separated values (CSV) format has \emph{14 features}
and has 0.9\% missing values. The target variable is the \emph{income} that 
can be either \emph{more} or \emph{less or equal} than \$50000.

Due to the census's scope, there was no allocation of participants into 
experimental groups or need for randomization typical of controlled 
experiments. The dataset naturally includes a broad demographic diversity 
reflective of the national census effort.

\subsection{Procedure}

\subsubsection{Hypothesis}

As an initial analysis, we aim to explore how social factors such as gender
and race affect income levels. We want to understand if a specific gender 
is more likely to earn more.

We also intend to investigate how other factors interact with gender. One 
interesting analysis could be to examine how ethnic groups might experience
advantages or disadvantages combined with gender in terms of income disparity.

Another intriguing hypothesis would consider the income progression 
influenced by combinations of age and gender.

\subsubsection{Exploration}

\subsubsection{Features Description}

First, we need to conduct a thorough exploration of the dataset to understand
the underlying structure and quality of the data. This will involve assessing 
various features for their relevance and potential impact on the 
study's outcomes, particularly how they relate to income levels.
We will also present a summary in Table X.

```{r, echo=TRUE}
census <- read.csv("../adult.csv", na.strings =  c("?"))
head(census)
```



```{r, echo=TRUE}
missing_values <- census %>% summarise(across(everything(), ~sum(is.na(.))))
missing_values_ft <- missing_values %>%
  pivot_longer(cols = everything(), names_to = "Feature", values_to = "MissingValues")

ggplot(missing_values_ft, aes(x = Feature, y = MissingValues, fill = Feature)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Missing Values Per Feature", x = "Feature", y = "Missing Values") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") + 
  theme(plot.title = element_text(hjust = 0.5)) 
```

We get the following resume for the categories:





\subsection{Scoring (Optional)}
